\textbf{Seasonal Dynamics of Epiphytic Microbial Communities on Marine Macrophyte Surfaces}

### 2020-04-19 submission/manuscript.tex
References optimization.

Sequencing of the 16S rRNA V4 region yielded a total of 1.8 million
sequences after quality curation and exclusion of eukaryotic,
chloroplast, mitochondrial and no relative sequences
(\autoref{supp-nseq_notus}). A total of 35 samples originating from
epiphytic archaeal and bacterial communities associated with surfaces of
the seagrass \emph{C. nodosa} and macroalga \emph{C. cylindracea} were
analyzed. In addition, 18 samples (one of the samples was sequenced two
times) originating from picoplankton archaeal and bacterial communities
in the ambient seawater were also processed for comparison. The number
of reads per sample ranged between 8,409 and 77,463 sequences
(\autoref{supp-nseq_notus}). Even when the highest sequencing effort was
applied the rarefaction curves did not level off that is a common
observation in high-throughput 16S rRNA amplicon sequencing approaches
(\autoref{supp-rarefaction}). Following quality curation and exclusion
of sequences mentioned before reads were clustered into 28,702 different
OTUs at a similarity level of 97 \si{\percent}. Read numbers were
normalized to the minimum number of sequences, 8,409
(\autoref{supp-nseq_notus}), through rarefaction resulting in 17,055
different OTUs that ranged from 370 to 2,033 OTUs per sample
(\autoref{supp-calculators}). To determine seasonal changes of richness
and diversity the Observed Number of OTUs, Chao1, ACE, Exponential
Shannon (Jost, 2006) and Inverse Simpson were calculated after
normalization through rarefaction. Generally, richness estimators and
diversity indices showed similar trends. On average, higher values were
found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,691.5 ±
144.4 OTUs{]} and monospecific {[}Number of OTUs, 1,733.3 ± 156.8
OTUs{]}), middle values for \emph{C. nodosa} (Number of OTUs, 1,060.9 ±
216.1 OTUs) and lower values for picoplankton communities in the ambient
seawater (Number of OTUs, 527.5 ± 147.8 OTUs)
(\autoref{supp-calculators}). Seasonal changes did not show such large
dissimilarities. \emph{C. nodosa} communities showed a slow increase
towards the end of the study, while \emph{C. cylindracea} (mixed and
monospecific) communities were characterized by slightly larger values
in Spring and Summer in comparison to Autumn and Winter
(\autoref{supp-calculators}).

To determine the proportion of shared archaeal and bacterial OTUs and
communities sampled in different environments the Jaccard's Similarity
Coefficient on presence-absence data and Bray-Curtis Similarity
Coefficient were, respectively, calculated. Coefficients were determined
after normalization through rarefaction and binning of samples from a
particular environment. The highest proportion of shared OTUs and
community was found between mixed and monospecific \emph{C. cylindracea}
(Jaccard, 0.35; Bray-Curtis, 0.77), while lower shared values were
calculated between seawater and epiphytic communities
(\autoref{matrix}). Shared proportion between \emph{C. nodosa} and
\emph{C. cylindracea} were approximately in the middle between these two
extremes. To assess seasonal changes in the proportion of shared OTUs
and communities the Jaccard's and Bray-Curtis Similarity Coefficients
were calculated between consecutive sampling points (\autoref{shared}).
Both coefficients showed similar trends. Temporal proportional changes
were more pronounced for seawater in comparison to \emph{C. nodosa} and
especially \emph{C. cylindracea} associated communities
(\autoref{shared}). In addition, only 0.4 -- 1.0 \si{\percent} of OTUs
from each surface associated community were found at every time point.
These OTUs also made a high proportion of total sequences (40.3 -- 53.7
\si{\percent}). To further disentangle the environmental and seasonal
community dissimilarity a Principal Coordinates Analysis (PCoA) based on
Bray-Curtis distances and OTU abundances was applied. It showed a clear
separation between planktonic and surface associated communities
(\autoref{pcoa}). In addition, a separation of epiphytic bacterial and
archaeal communities based on host species was determined. This
separation was further supported by ANOSIM (R = 0.96, \emph{p}
\textless{} 0.001). Seasonal changes of \emph{C. nodosa} associated
communities indicated a separation between Spring, Summer and
Autumn/Winter samples (ANOSIM, R = 0.54, \emph{p} \textless{} 0.01). For
\emph{C. cylindracea} associated communities a separation between Summer
and Autumn/Winter/Spring samples was observed that was not so strongly
supported (ANOSIM, R = 0.31, \emph{p} \textless{} 0.05)
(\autoref{pcoa}).

@@

Seasonal richness changes in the epiphytic community were substantial as
could be observed in the proportion of OTUs that could be found at every
sampling time (\(\leq\) 1.0 \si{\percent}). Interestingly, these OTUs
were accounting for a high proportion of sequences (\(\leq\) 53.7
\si{\percent}). A very similar proportion of persistent OTUs and their
sequence contribution was reported in high frequency studies describing
seasonal picoplankton changes (Gilbert \emph{et al.}, 2009, 2012). In
comparison to the seawater community, a lower degree of seasonal shifts
was observed for the surface associated communities. It seems,
microniches on the surfaces of macrophytes are providing more stable
conditions in comparison to the seawater. At the level of OTUs seasonal
changes of \emph{C. nodosa} and \emph{C. cylindracea} associated
communities were identified that could be linked to the growth cycle of
the seagrass and macroalgae (M. Najdek, personal communication).
\emph{C. nodosa} was characterized by a Spring community during maximum
seagrass proliferation, a Summer community during a biomass maximum and
a Autumn/Winter community during a biomass senescence. In contrast,
\emph{C. cylindracea} started to proliferate in late Spring and was
characterized only by a Summer community during maximal biomass increase
and by a Autumn/Winter/Spring community when the biomass was at the peak
and the settlement started to subsequently decay. Similar seasonal
changes in the epiphytic community was also described for other
macroalgae (Tujula \emph{et al.}, 2010; Lachnit \emph{et al.}, 2011).
Higher temporal stability of \emph{C. cylindracea} surface communities
in comparison to \emph{C. nodosa} were also observed in the higher
proportion of shared communities between two consecutive sampling
points.

@@

Obtained sequences were analyzed on the computer cluster Isabella
(University Computing Center, University of Zagreb) using mothur
(version 1.43.0) (Schloss \emph{et al.}, 2009) according to the MiSeq
Standard Operating Procedure (MiSeq SOP;
\url{https://mothur.org/wiki/MiSeq_SOP}) (Kozich \emph{et al.}, 2013)
and recommendations given from the Riffomonas project to enhance data
reproducibility (\url{http://www.riffomonas.org/}). For alignment and
classification of sequences the SILVA SSU Ref NR 99 database (release
138; \url{http://www.arb-silva.de}) was used (Quast \emph{et al.}, 2013;
Yilmaz \emph{et al.}, 2014). Pipeline data processing and visualization
was done using R (version 3.6.0) (R Core Team, 2019), packages vegan
(version 2.5-6) (Oksanen \emph{et al.}, 2019), and tidyverse (version
1.3.0) (Wickham \emph{et al.}, 2019) and multiple other packages (Xie,
2014, 2015, 2020; Neuwirth, 2014; Xie \emph{et al.}, 2018; Y. Xie,
2019a, 2019b; Allaire \emph{et al.}, 2019; Zhu, 2019). The detailed
analysis procedure including the R Markdown file for this paper are
available as a GitHub repository
(\url{https://github.com/mkorlevic/Korlevic_EpiphyticDynamics_EnvironMicrobiol_2020}).
Based on the ATCC MSA-1002 mock community included in the analysis an
average sequencing error rate of 0.01 \si{\percent} was determined,
which is in line with previously reported values for next-generation
sequencing data (Kozich \emph{et al.}, 2013; Schloss \emph{et al.},
2016). In addition, the negative controls processed together with the
samples yielded on average only 2 sequences after sequence quality
curation.

@@

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Abraham2014}{}%
Abraham, W.R. and Rohde, M. (2014) The family \emph{Hyphomonadaceae}. In
\emph{The prokaryotes: Alphaproteobacteria and Betaproteobacteria}.
Rosenberg, E., DeLong, E.F., Lory, S., Stackebrandt, E., and Thompson,
F. (eds). Berlin, Heidelberg: Springer-Verlag, pp. 283--299.

\leavevmode\hypertarget{ref-Allaire2019}{}%
Allaire, J.J., Xie, Y., McPherson, J., Luraschi, J., Ushey, K., Atkins,
A., et al. (2019) Rmarkdown: dynamic documents for R.

@@

\leavevmode\hypertarget{ref-Bengtsson2010}{}%
Bengtsson, M., Sjøtun, K., and Øvreås, L. (2010) Seasonal dynamics of
bacterial biofilms on the kelp \emph{Laminaria hyperborea}. \emph{Aquat
Microb Ecol} \textbf{60}: 71--83.

@@

\leavevmode\hypertarget{ref-Burke2011a}{}%
Burke, C., Steinberg, P., Rusch, D., Kjelleberg, S., and Thomas, T.
(2011) Bacterial community assembly based on functional genes rather
than species. \emph{Proc Natl Acad Sci U S A} \textbf{108}:
14288--14293.

\leavevmode\hypertarget{ref-Burke2011b}{}%
Burke, C., Thomas, T., Lewis, M., Steinberg, P., and Kjelleberg, S.
(2011) Composition, uniqueness and variability of the epiphytic
bacterial community of the green alga \emph{Ulva australis}. \emph{ISME
J} \textbf{5}: 590--600.

@@

\leavevmode\hypertarget{ref-Egan2013}{}%
Egan, S., Harder, T., Burke, C., Steinberg, P., Kjelleberg, S., and
Thomas, T. (2013) The seaweed holobiont: understanding seaweed-bacteria
interactions. \emph{FEMS Microbiol Rev} \textbf{37}: 462--476.

@@

\leavevmode\hypertarget{ref-Hollants2013}{}%
Hollants, J., Leliaert, F., De Clerck, O., and Willems, A. (2013) What
we can learn from sushi: a review on seaweed-bacterial associations.
\textbf{83}: 1--16.

@@

\leavevmode\hypertarget{ref-Kuever2014}{}%
Kuever, J. (2014) The family \emph{Desulfobulbaceae}. In \emph{The
prokaryotes: Deltaproteobacteria and Epsilonproteobacteria}. Rosenberg,
E., DeLong, E.F., Lory, S., Stackebrandt, E., and Thompson, F. (eds).
Berlin, Heidelberg: Springer-Verlag, pp. 75--86.

\leavevmode\hypertarget{ref-Lachnit2009}{}%
Lachnit, T., Blümel, M., Imhoff, J.F., and Wahl, M. (2009) Specific
epibacterial communities on macroalgae: phylogeny matters more than
habitat. \emph{Aquat Biol} \textbf{5}: 181--186.

@@

\leavevmode\hypertarget{ref-Margulis1991}{}%
Margulis, L. (1991) Symbiogenesis and symbionticism. In \emph{Symbiosis
as a source of evolutionary innovation: Speciation and morphogenesis}.
Margulis, L. and Fester, R. (eds). Cambridge, Massachusetts: The MIT
Press, pp. 1--14.

@@

\leavevmode\hypertarget{ref-McIlroy2014}{}%
McIlroy, S.J. and Nielsen, P.H. (2014) The family \emph{Saprospiraceae}.
In \emph{The prokaryotes: Other major lineages of Bacteria and the
Archaea}. Rosenberg, E., DeLong, E.F., Lory, S., Stackebrandt, E., and
Thompson, F. (eds). Berlin, Heidelberg: Springer-Verlag, pp. 863--889.

\leavevmode\hypertarget{ref-Michelou2013}{}%
Michelou, V.K., Caporaso, J.G., Knight, R., and Palumbi, S.R. (2013) The
ecology of microbial communities associated with \emph{Macrocystis
pyrifera}. \emph{PLoS One} \textbf{8}: e67480.

\leavevmode\hypertarget{ref-Miranda2013}{}%
Miranda, L.N., Hutchison, K., Grossman, A.R., and Brawley, S.H. (2013)
Diversity and abundance of the bacterial community of the red macroalga
\emph{Porphyra umbilicalis}: did bacterial farmers produce macroalgae?
\emph{PLoS One} \textbf{8}: e58269.

@@

\leavevmode\hypertarget{ref-Oksanen2019}{}%
Oksanen, J., Blanchet, F.G., Friendly, M., Kindt, R., Legendre, P.,
McGlinn, D., et al. (2019) Vegan: community ecology package.

\leavevmode\hypertarget{ref-Parada2016}{}%
Parada, A.E., Needham, D.M., and Fuhrman, J.A. (2016) Every base
matters: assessing small subunit rRNA primers for marine microbiomes
with mock communities, time series and global field samples.
\emph{Environ Microbiol} \textbf{18}: 1403--1414.

\leavevmode\hypertarget{ref-Pujalte2014}{}%
Pujalte, M.J., Lucena, T., Ruvira, M.A., Arahal, D.R., and Macián, M.C.
(2014) The family \emph{Rhodobacteraceae}. In \emph{The prokaryotes:
Alphaproteobacteria and Betaproteobacteria}. Rosenberg, E., DeLong,
E.F., Lory, S., Stackebrandt, E., and Thompson, F. (eds). Berlin,
Heidelberg: Springer-Verlag, pp. 439--512.

\leavevmode\hypertarget{ref-Quast2013}{}%
Quast, C., Pruesse, E., Yilmaz, P., Gerken, J., Schweer, T., Yarza, P.,
et al. (2013) The SILVA ribosomal RNA gene database project: improved
data processing and web-based tools. \emph{Nucleic Acids Res}
\textbf{41}: D590--D596.

\leavevmode\hypertarget{ref-RCoreTeam2019}{}%
R Core Team (2019) A language and environment for statistical computing.

\leavevmode\hypertarget{ref-Reisser2014}{}%
Reisser, J., Shaw, J., Hallegraeff, G., Proietti, M., Barnes, D.K.A.,
Thums, M., et al. (2014) Millimeter-sized marine plastics: a new pelagic
habitat for microorganisms and invertebrates. \emph{PLoS One}
\textbf{9}: e100289.

@@

\leavevmode\hypertarget{ref-Rickert2016}{}%
Rickert, E., Wahl, M., Link, H., Richter, H., and Pohnert, G. (2016)
Seasonal variations in surface metabolite composition of \emph{Fucus
vesiculosus} and \emph{Fucus serratus} from the Baltic Sea. \emph{PLoS
One} \textbf{11}: e0168196.

@@

\leavevmode\hypertarget{ref-Salaun2012}{}%
Salaün, S., La Barre, S., Santos-Goncalvez, M.D., Potin, P., Haras, D.,
and Bazire, A. (2012) Influence of exudates of the kelp \emph{Laminaria
digitata} on biofilm formation of associated and exogenous bacterial
epiphytes. \emph{Microb Ecol} \textbf{64}: 359--369.

\leavevmode\hypertarget{ref-Schloss2016a}{}%
Schloss, P.D., Jenior, M.L., Koumpouras, C.C., Westcott, S.L., and
Highlander, S.K. (2016) Sequencing 16S rRNA gene fragments using the
PacBio SMRT DNA sequencing system. \emph{PeerJ} \textbf{4}: e1869.

\leavevmode\hypertarget{ref-Schloss2009}{}%
Schloss, P.D., Westcott, S.L., Ryabin, T., Hall, J.R., Hartmann, M.,
Hollister, E.B., et al. (2009) Introducing mothur: open-source,
platform-independent, community-supported software for describing and
comparing microbial communities. \emph{Appl Environ Microbiol}
\textbf{75}: 7537--7541.

\leavevmode\hypertarget{ref-Tarquinio2019}{}%
Tarquinio, F., Hyndes, G.A., Laverock, B., Koenders, A., and Säwström,
C. (2019) The seagrass holobiont: understanding seagrass-bacteria
interactions and their role in seagrass ecosystem functioning.
\emph{FEMS Microbiol Lett} \textbf{366}: fnz057.

@@

\leavevmode\hypertarget{ref-Xie2015}{}%
Xie, Y. (2015) Dynamic documents with R and knitr, 2nd ed. Boca Raton,
Florida: Chapman; Hall/CRC.

\leavevmode\hypertarget{ref-Xie2014}{}%
Xie, Y. (2014) Knitr: a comprehensive tool for reproducible research in
R. In \emph{Implementing reproducible computational research}. Stodden,
V., Leisch, F., and Peng, R.D. (eds). New York: Chapman; Hall/CRC, pp.
3--32.

\leavevmode\hypertarget{ref-Xie2019a}{}%
Xie, Y. (2019a) Knitr: a general-purpose package for dynamic report
generation in R.

\leavevmode\hypertarget{ref-Xie2019}{}%
Xie, Y. (2019b) TinyTeX: a lightweight, cross-platform, and
easy-to-maintain LaTeX distribution based on TeX Live. \emph{TUGboat}
\textbf{40}: 30--32.

\leavevmode\hypertarget{ref-Xie2020}{}%
Xie, Y. (2020) TinyTex: helper functions to install and maintain 'TeX
Live', and compile 'LaTeX' documents.

\leavevmode\hypertarget{ref-Xie2018}{}%
Xie, Y., Allaire, J.J., and Grolemund, G. (2018) R Markdown: the
definitive guide, 1st ed. Boca Raton, Florida: Chapman; Hall/CRC.

\leavevmode\hypertarget{ref-Xie2017}{}%
Xie, Z., Lin, W., and Luo, J. (2017) Comparative phenotype and genome
analysis of \emph{Cellvibrio} sp. PR1, a xylanolytic and agarolytic
bacterium from the Pearl River. \emph{Biomed Res Int} \textbf{2017}:
6304248.

@@

\leavevmode\hypertarget{ref-Zhu2019}{}%
Zhu, H. (2019) KableExtra: construct complex table with 'kable' and pipe
syntax.

>>>>

Sequencing of the 16S rRNA V4 region yielded a total of 1.8 million
sequences after quality curation and exclusion of eukaryotic,
chloroplast, mitochondrial and no relative sequences
(\autoref{supp-nseq_notus}). A total of 35 samples originating from
epiphytic archaeal and bacterial communities associated with surfaces of
the seagrass \emph{C. nodosa} and macroalga \emph{C. cylindracea} were
analyzed. In addition, 18 samples (one of the samples was sequenced two
times) originating from picoplankton archaeal and bacterial communities
in the ambient seawater were also processed for comparison. The number
of reads per sample ranged between 8,408 and 77,463 sequences
(\autoref{supp-nseq_notus}). Even when the highest sequencing effort was
applied the rarefaction curves did not level off that is a common
observation in high-throughput 16S rRNA amplicon sequencing approaches
(\autoref{supp-rarefaction}). Following quality curation and exclusion
of sequences mentioned before reads were clustered into 28,750 different
OTUs at a similarity level of 97 \si{\percent}. Read numbers were
normalized to the minimum number of sequences, 8,408
(\autoref{supp-nseq_notus}), through rarefaction resulting in 17,115
different OTUs that ranged from 372 to 2,084 OTUs per sample
(\autoref{supp-calculators}). To determine seasonal changes of richness
and diversity the Observed Number of OTUs, Chao1, ACE, Exponential
Shannon (Jost, 2006) and Inverse Simpson were calculated after
normalization through rarefaction. Generally, richness estimators and
diversity indices showed similar trends. On average, higher values were
found for \emph{C. cylindracea} (mixed {[}Number of OTUs, 1,684.6 ±
140.0 OTUs{]} and monospecific {[}Number of OTUs, 1,735 ± 172.0
OTUs{]}), middle values for \emph{C. nodosa} (Number of OTUs, 1,059.7 ±
208.8 OTUs) and lower values for picoplankton communities in the ambient
seawater (Number of OTUs, 531.7 ± 141.4 OTUs)
(\autoref{supp-calculators}). Seasonal changes did not show such large
dissimilarities. \emph{C. nodosa} communities showed a slow increase
towards the end of the study, while \emph{C. cylindracea} (mixed and
monospecific) communities were characterized by slightly larger values
in Spring and Summer in comparison to Autumn and Winter
(\autoref{supp-calculators}).

To determine the proportion of shared archaeal and bacterial OTUs and
communities sampled in different environments the Jaccard's Similarity
Coefficient on presence-absence data and Bray-Curtis Similarity
Coefficient were, respectively, calculated. Coefficients were determined
after normalization through rarefaction and binning of samples from a
particular environment. The highest proportion of shared OTUs and
community was found between mixed and monospecific \emph{C. cylindracea}
(Jaccard, 0.35; Bray-Curtis, 0.77), while lower shared values were
calculated between seawater and epiphytic communities
(\autoref{matrix}). Shared proportion between \emph{C. nodosa} and
\emph{C. cylindracea} were approximately in the middle between these two
extremes. To assess seasonal changes in the proportion of shared OTUs
and communities the Jaccard's and Bray-Curtis Similarity Coefficients
were calculated between consecutive sampling points (\autoref{shared}).
Both coefficients showed similar trends. Temporal proportional changes
were more pronounced for seawater in comparison to \emph{C. nodosa} and
especially \emph{C. cylindracea} associated communities
(\autoref{shared}). In addition, only 0.4 -- 1.0 \si{\percent} of OTUs
from each surface associated community were found at every time point.
These OTUs also made a high proportion of total sequences (40.1 -- 51.4
\si{\percent}). To further disentangle the environmental and seasonal
community dissimilarity a Principal Coordinates Analysis (PCoA) based on
Bray-Curtis distances and OTU abundances was applied. It showed a clear
separation between planktonic and surface associated communities
(\autoref{pcoa}). In addition, a separation of epiphytic bacterial and
archaeal communities based on host species was determined. This
separation was further supported by ANOSIM (R = 0.96, \emph{p}
\textless{} 0.001). Seasonal changes of \emph{C. nodosa} associated
communities indicated a separation between Spring, Summer and
Autumn/Winter samples (ANOSIM, R = 0.54, \emph{p} \textless{} 0.01). For
\emph{C. cylindracea} associated communities a separation between Summer
and Autumn/Winter/Spring samples was observed that was not so strongly
supported (ANOSIM, R = 0.32, \emph{p} \textless{} 0.01)
(\autoref{pcoa}).

@@

Seasonal richness changes in the epiphytic community were substantial as
could be observed in the proportion of OTUs that could be found at every
sampling time (\(\leq\) 1.0 \si{\percent}). Interestingly, these OTUs
were accounting for a high proportion of sequences (\(\leq\) 51.4
\si{\percent}). A very similar proportion of persistent OTUs and their
sequence contribution was reported in high frequency studies describing
seasonal picoplankton changes (Gilbert \emph{et al.}, 2009, 2012). In
comparison to the seawater community, a lower degree of seasonal shifts
was observed for the surface associated communities. It seems,
microniches on the surfaces of macrophytes are providing more stable
conditions in comparison to the seawater. At the level of OTUs seasonal
changes of \emph{C. nodosa} and \emph{C. cylindracea} associated
communities were identified that could be linked to the growth cycle of
the seagrass and macroalgae (M. Najdek, personal communication).
\emph{C. nodosa} was characterized by a Spring community during maximum
seagrass proliferation, a Summer community during a biomass maximum and
a Autumn/Winter community during a biomass senescence. In contrast,
\emph{C. cylindracea} started to proliferate in late Spring and was
characterized only by a Summer community during maximal biomass increase
and by a Autumn/Winter/Spring community when the biomass was at the peak
and the settlement started to subsequently decay. Similar seasonal
changes in the epiphytic community was also described for other
macroalgae (Tujula \emph{et al.}, 2010; Lachnit \emph{et al.}, 2011).
Higher temporal stability of \emph{C. cylindracea} surface communities
in comparison to \emph{C. nodosa} were also observed in the higher
proportion of shared communities between two consecutive sampling
points.

@@

Obtained sequences were analyzed on the computer cluster Isabella
(University Computing Center, University of Zagreb) using mothur
(version 1.43.0) (Schloss \emph{et al.}, 2009) according to the MiSeq
Standard Operating Procedure (MiSeq SOP;
\url{https://mothur.org/wiki/MiSeq_SOP}) (Kozich \emph{et al.}, 2013)
and recommendations given from the Riffomonas project to enhance data
reproducibility (\url{http://www.riffomonas.org/}). For alignment and
classification of sequences the SILVA SSU Ref NR 99 database (release
138; \url{http://www.arb-silva.de}) was used (Quast \emph{et al.}, 2013;
Yilmaz \emph{et al.}, 2014). Pipeline data processing and visualization
was done using R (version 3.6.0) (R Core Team, 2019), packages vegan
(version 2.5-6) (Oksanen \emph{et al.}, 2019), and tidyverse (version
1.3.0) (Wickham \emph{et al.}, 2019) and multiple other packages (Xie,
2014, 2015, 2020; Neuwirth, 2014; Xie \emph{et al.}, 2018; Y. Xie,
2019b, 2019a; Allaire \emph{et al.}, 2019; Zhu, 2019). The detailed
analysis procedure including the R Markdown file for this paper are
available as a GitHub repository
(\url{https://github.com/mkorlevic/Korlevic_EpiphyticDynamics_EnvironMicrobiol_2020}).
Based on the ATCC MSA-1002 mock community included in the analysis an
average sequencing error rate of 0.01 \si{\percent} was determined,
which is in line with previously reported values for next-generation
sequencing data (Kozich \emph{et al.}, 2013; Schloss \emph{et al.},
2016). In addition, the negative controls processed together with the
samples yielded on average only 2 sequences after sequence quality
curation.

@@

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Abraham2014}{}%
Abraham, W.R. and Rohde, M. (2014) The family \emph{Hyphomonadaceae}. In
\emph{The Prokaryotes: Alphaproteobacteria and Betaproteobacteria}.
Rosenberg, E., DeLong, E.F., Lory, S., Stackebrandt, E., and Thompson,
F. (eds). Berlin, Heidelberg: Springer-Verlag, pp. 283--299.

\leavevmode\hypertarget{ref-Allaire2019}{}%
Allaire, J.J., Xie, Y., McPherson, J., Luraschi, J., Ushey, K., Atkins,
A., et al. (2019) Rmarkdown: Dynamic documents for R.

@@

\leavevmode\hypertarget{ref-Bengtsson2010}{}%
Bengtsson, M., Sjøtun, K., and Øvre\textbackslash{}aas, L. (2010)
Seasonal dynamics of bacterial biofilms on the kelp \emph{Laminaria
hyperborea}. \emph{Aquat Microb Ecol} \textbf{60}: 71--83.

@@

\leavevmode\hypertarget{ref-Burke2011}{}%
Burke, C., Steinberg, P., Rusch, D., Kjelleberg, S., and Thomas, T.
(2011) Bacterial community assembly based on functional genes rather
than species. \emph{Proc Natl Acad Sci U S A} \textbf{108}:
14288--14293.

\leavevmode\hypertarget{ref-Burke2011a}{}%
Burke, C., Thomas, T., Lewis, M., Steinberg, P., and Kjelleberg, S.
(2011) Composition, uniqueness and variability of the epiphytic
bacterial community of the green alga \emph{Ulva australis}. \emph{ISME
J} \textbf{5}: 590--600.

@@

\leavevmode\hypertarget{ref-Egan2013}{}%
Egan, S., Harder, T., Burke, C., Steinberg, P., Kjelleberg, S., and
Thomas, T. (2013) The seaweed holobiont: Understanding seaweed-bacteria
interactions. \emph{FEMS Microbiol Rev} \textbf{37}: 462--476.

@@

\leavevmode\hypertarget{ref-Hollants2013}{}%
Hollants, J., Leliaert, F., De Clerck, O., and Willems, A. (2013) What
we can learn from sushi: A review on seaweed-bacterial associations.
\emph{FEMS Microbiol Ecol} \textbf{83}:

@@

\leavevmode\hypertarget{ref-Kuever2014}{}%
Kuever, J. (2014) The family \emph{Desulfobulbaceae}. In \emph{The
Prokaryotes: Deltaproteobacteria and Epsilonproteobacteria}. Rosenberg,
E., DeLong, E.F., Lory, S., Stackebrandt, E., and Thompson, F. (eds).
Berlin, Heidelberg: Springer-Verlag, pp. 75--86.

\leavevmode\hypertarget{ref-Lachnit2009}{}%
Lachnit, T., Blümel, M., Imhoff, J.F., and Wahl, M. (2009) Specific
epibacterial communities on macroalgae: Phylogeny matters more than
habitat. \emph{Aquat Biol} \textbf{5}: 181--186.

@@

\leavevmode\hypertarget{ref-Margulis1991}{}%
Margulis, L. (1991) Symbiogenesis and symbionticism. In \emph{Symbiosis
as a Source of Evolutionary Innovation: Speciation and Morphogenesis}.
Margulis, L. and Fester, R. (eds). Cambridge, Massachusetts: The MIT
Press, pp. 1--14.

@@

\leavevmode\hypertarget{ref-McIlroy2014}{}%
McIlroy, S.J. and Nielsen, P.H. (2014) The family \emph{Saprospiraceae}.
In \emph{The Prokaryotes: Other Major Lineages of Bacteria and the
Archaea}. Rosenberg, E., DeLong, E.F., Lory, S., Stackebrandt, E., and
Thompson, F. (eds). Berlin, Heidelberg: Springer-Verlag, pp. 863--889.

\leavevmode\hypertarget{ref-Michelou2013}{}%
Michelou, V.K., Caporaso, J.G., Knight, R., and Palumbi, S.R. (2013) The
ecology of microbial communities associated with \emph{Macrocystis
pyrifera}. \emph{PloS One} \textbf{8}: e67480.

\leavevmode\hypertarget{ref-Miranda2013}{}%
Miranda, L.N., Hutchison, K., Grossman, A.R., and Brawley, S.H. (2013)
Diversity and abundance of the bacterial community of the red macroalga
\emph{Porphyra umbilicalis}: Did bacterial farmers produce macroalgae?
\emph{PloS One} \textbf{8}: e58269.

@@

\leavevmode\hypertarget{ref-Oksanen2019}{}%
Oksanen, J., Blanchet, F.G., Friendly, M., Kindt, R., Legendre, P.,
McGlinn, D., et al. (2019) Vegan: Community ecology package.

\leavevmode\hypertarget{ref-Parada2016}{}%
Parada, A.E., Needham, D.M., and Fuhrman, J.A. (2016) Every base
matters: Assessing small subunit rRNA primers for marine microbiomes
with mock communities, time series and global field samples.
\emph{Environ Microbiol} \textbf{18}: 1403--1414.

\leavevmode\hypertarget{ref-Pujalte2014}{}%
Pujalte, M.J., Lucena, T., Ruvira, M.A., Arahal, D.R., and Macián, M.C.
(2014) The family \emph{Rhodobacteraceae}. In \emph{The Prokaryotes:
Alphaproteobacteria and Betaproteobacteria}. Rosenberg, E., DeLong,
E.F., Lory, S., Stackebrandt, E., and Thompson, F. (eds). Berlin,
Heidelberg: Springer-Verlag, pp. 439--512.

\leavevmode\hypertarget{ref-Quast2013}{}%
Quast, C., Pruesse, E., Yilmaz, P., Gerken, J., Schweer, T., Yarza, P.,
et al. (2013) The SILVA ribosomal RNA gene database project: Improved
data processing and web-based tools. \emph{Nucleic Acids Res}
\textbf{41}: D590--D596.

\leavevmode\hypertarget{ref-RCoreTeam2019}{}%
R Core Team (2019) A language and environment for statistical computing,
Vienna, Austria: R Foundation for Statistical Computing.

\leavevmode\hypertarget{ref-Reisser2014}{}%
Reisser, J., Shaw, J., Hallegraeff, G., Proietti, M., Barnes, D.K.A.,
Thums, M., et al. (2014) Millimeter-sized marine plastics: A new pelagic
habitat for microorganisms and invertebrates. \emph{PloS One}
\textbf{9}: e100289.

@@

\leavevmode\hypertarget{ref-Rickert2016}{}%
Rickert, E., Wahl, M., Link, H., Richter, H., and Pohnert, G. (2016)
Seasonal variations in surface metabolite composition of \emph{Fucus
vesiculosus} and \emph{Fucus serratus} from the Baltic Sea. \emph{PloS
One} \textbf{11}: e0168196.

@@

\leavevmode\hypertarget{ref-Salaun2012}{}%
Salaün, S., La Barre, S., Santos-Goncalvez, M.D., Potin, P., Haras, D.,
and Bazire, A. (2012) Influence of exudates of the kelp Laminaria
digitata on biofilm formation of associated and exogenous bacterial
epiphytes. \emph{Microb Ecol} \textbf{64}: 359--369.

\leavevmode\hypertarget{ref-Schloss2016}{}%
Schloss, P.D., Jenior, M.L., Koumpouras, C.C., Westcott, S.L., and
Highlander, S.K. (2016) Sequencing 16S rRNA gene fragments using the
PacBio SMRT DNA sequencing system. \emph{PeerJ} \textbf{4}: e1869.

\leavevmode\hypertarget{ref-Schloss2009}{}%
Schloss, P.D., Westcott, S.L., Ryabin, T., Hall, J.R., Hartmann, M.,
Hollister, E.B., et al. (2009) Introducing mothur: Open-source,
platform-independent, community-supported software for describing and
comparing microbial communities. \emph{Appl Environ Microbiol}
\textbf{75}: 7537--7541.

\leavevmode\hypertarget{ref-Tarquinio2019}{}%
Tarquinio, F., Hyndes, G.A., Laverock, B., Koenders, A., and Säwström,
C. (2019) The seagrass holobiont: Understanding seagrass-bacteria
interactions and their role in seagrass ecosystem functioning.
\emph{FEMS Microbiol Lett} \textbf{366}: fnz057.

@@

\leavevmode\hypertarget{ref-Xie2015}{}%
Xie, Y. (2015) Dynamic Documents with R and knitr, 2nd ed. Boca Raton,
Florida: Chapman and Hall/CRC.

\leavevmode\hypertarget{ref-Xie2014}{}%
Xie, Y. (2014) Knitr: A comprehensive tool for reproducible research in
R. In \emph{Implementing Reproducible Computational Research}. Stodden,
V., Leisch, F., and Peng, R.D. (eds). New York: Chapman and Hall/CRC,
pp. 3--32.

\leavevmode\hypertarget{ref-Xie2019}{}%
Xie, Y. (2019a) Knitr: A general-purpose package for dynamic report
generation in R.

\leavevmode\hypertarget{ref-Xie2019a}{}%
Xie, Y. (2019b) TinyTeX: A lightweight, cross-platform, and
easy-to-maintain LaTeX distribution based on TeX Live. \emph{TUGboat}
\textbf{40}: 30--32.

\leavevmode\hypertarget{ref-Xie2020}{}%
Xie, Y. (2020) TinyTex: Helper functions to install and maintain 'TeX
Live', and compile 'LaTeX' documents.

\leavevmode\hypertarget{ref-Xie2018}{}%
Xie, Y., Allaire, J.J., and Grolemund, G. (2018) R Markdown: The
Definitive Guide, 1st ed. Boca Raton, Florida: Chapman and Hall/CRC.

\leavevmode\hypertarget{ref-Xie2017}{}%
Xie, Z., Lin, W., and Luo, J. (2017) Comparative phenotype and genome
analysis of \emph{Cellvibrio} sp. PR1, a xylanolytic and agarolytic
bacterium from the Pearl River. \emph{BioMed Res Int} \textbf{2017}:
6304248.

@@

\leavevmode\hypertarget{ref-Zhu2019}{}%
Zhu, H. (2019) KableExtra: Construct complex table with 'kable' and pipe
syntax.
